External priors for the next generation of CMB experiments.

### 2015-09-11 Notes/N_eff_prior_paper.tex
last section with new suggestions for the figures

% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
%  Add 'draft' option to mark overfull boxes with black boxes
%  Add 'showpacs' option to make PACS codes appear
%  Add 'showkeys' option to make keywords appear
\documentclass[aps,prd,reprint,superscriptaddress]{revtex4-1}
%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4-1}
%\documentclass[aps,prl,reprint,groupedaddress]{revtex4-1}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{lineno}
\usepackage{graphicx}
%\input{papers_packages_command._include_tex}
\newcommand\refsec[1]{\S\ref{sec:#1}}
\newcommand\refeq[1]{Eq.~(\ref{eqn:#1})}
\newcommand{\refssec}[1]{section~\ref{subsec:#1}}
\newcommand{\reffig}[1]{Fig.~\ref{fig:#1}}
% For various journals
%\newcommand{\apj}{ApJ}
\newcommand{\physrep}{Physics Reports}
\newcommand{\aapr}{A\&A Rev.}
\newcommand{\apjl}{ApJL}
\newcommand{\apjs}{ApJS}
\newcommand{\aap}{A\&A}
\newcommand{\mnras}{MNRAS}
%\newcommand{\prd}{Phys. Rev. D}
%\newcommand{\physrev}{Phys. Rev.}
\newcommand{\physrevlett}{Phys. Rev. Lett.}
\newcommand{\jcap}{{\em JCAP }}
\newcommand{\aj}{AJ }

\bibliographystyle{apsrev4-1}

@@

\section{Assumptions and methods \label{sec:methods}}
In this paper we want to measure quantitatively the effect of external priors on the  cosmological parameters constraints derived from S4 CMB experiments.
We start this section in \refssec{fm-formalism} by introducing the Fisher matrix formalism, a simple but powerful technique widely used to forecast future experimental constraints. To apply this technique, we need to specify our cosmological model ($\Lambda$CDM plus neutrinos and dark energy extensions in our case) as well as a set of fiducial parameters and the specifications of S4 CMB experiments. These will be presented in \refssec{cosmo-noise}.

@@

\section{Results \label{sec:results}}
Our main results are shown in \reffig{prior_massless_neutrinos} \reffig{prior_w}, \reffig{prior_re_optical_depth}

@@

\begin{table}[htdp]

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics{prior_massless_neutrinos_snow_mass.pdf}
\caption{$N_{eff}$ costraints mainly benefit from $\Omega_{b}$ and $n_{s}$ external priors.}
\label{fig:prior_massless_neutrinos}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics{prior_re_optical_depth_snow_mass.pdf}
\caption{$\tau$ constraints are limited by the well known degeneracies with $A_{s}$. An external prior on $A_{s}$ already tightly constrained from CMB alone can still be beneficial. }
\label{fig:prior_re_optical_depth}
\end{center}
\end{figure}

\section{Conclusions \label{sec:conclusions}}
--We find that...

--This means that supernova experiments CMB spatial distortions and large scale structure can/will/should

--Future works

>>>>

% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
%  Add 'draft' option to mark overfull boxes with black boxes
%  Add 'showpacs' option to make PACS codes appear
%  Add 'showkeys' option to make keywords appear
\documentclass[aps,prd,reprint,superscriptaddress]{revtex4-1}
%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4-1}
%\documentclass[aps,prl,reprint,groupedaddress]{revtex4-1}
%\usepackage{bm}
\usepackage{amsmath}
\usepackage{lineno}
\usepackage{graphicx}
%<<<<<<< HEAD
%\input{papers_packages_command._include_tex}

%\newcommand\refsec[1]{\S\ref{sec:#1}}
%\newcommand\refeq[1]{Eq.~(\ref{eqn:#1})}
%\input{papers_packages_command._include_tex}
\newcommand\refsec[1]{\S\ref{sec:#1}}
\newcommand\refeq[1]{Eq.~(\ref{eqn:#1})}
\newcommand{\refssec}[1]{section~\ref{subsec:#1}}
\newcommand{\reffig}[1]{Fig.~\ref{fig:#1}}
%\newcommand{\apj}{ApJ}
\newcommand{\physrep}{Physics Reports}
\newcommand{\aapr}{A\&A Rev.}
\newcommand{\apjl}{ApJL}
\newcommand{\apjs}{ApJS}
\newcommand{\aap}{A\&A}
\newcommand{\mnras}{MNRAS}
%\newcommand{\prd}{Phys. Rev. D}
\newcommand{\physrev}{Phys. Rev.}
\newcommand{\physrevlett}{Phys. Rev. Lett.}
\newcommand{\jcap}{{\em JCAP }}
\newcommand{\aj}{AJ }

%>>>>>>> 683231e77bc036bad411dd2244ca99280568d723
\bibliographystyle{apsrev4-1}

@@

\section{Assumptions and methods \label{sec:methods}}

{\bf A lot of this is standard and need not be included, which will make the paper much shorter: a good thing!}

In this paper we want to measure quantitatively the effect of external priors on the  cosmological parameters constraints derived from S4 CMB experiments.
We start this section in \refssec{fm-formalism} by introducing the Fisher matrix formalism, a simple but powerful technique widely used to forecast future experimental constraints. To apply this technique, we need to specify our cosmological model ($\Lambda$CDM plus neutrinos and dark energy extensions in our case) as well as a set of fiducial parameters and the specifications of S4 CMB experiments. These will be presented in \refssec{cosmo-noise}.

@@

\section{Results \label{sec:results}}

{\bf Let's open up one parameter at a time: first standard LCDM with neff, and then with mnu, and then with w. Also, let's show a curve in each plot with constraints on *all*  parameters, not just one of them, as that will give people a sense of what is gained if multiple priors can be obtained.}

\subsection{Relativistic Degrees of Freedom}

In the standard cosmology, there are three active neutrinos thermally produced in the early universe. Were they to decouple well before the epoch of electron-positron annihilation, their energy density after their decoupling would be equal to $3\times (7/8)\times (4/11)^{4/3}\,\rho_{\rm cmb}$, with the first factor capturing the contributions from the 3 active species; the second the difference between fermions and bosons; and the last the relative heating of the photons in the CMB by electron-positron annihilation. Decoupling is not a discrete event and occurs close to the time of electron-positron annihilation, so the neutrinos share a bit in the heating, with the factor of 3 replaced by $N_{\rm eff}=3.046$. The additional fraction depends not only on well-known neutrino scattering rates but also finite temperature quantum corrections. Were we to measure this deviation of $N_{\rm eff}$ from 3, it would be an indication that we understand well physics when the Universe was about a second old, so this is a natural target for upcoming experiments.

\begin{figure}[htbp]
\begin{center}
\includegraphics{prior_massless_neutrinos_snow_mass.pdf}
\caption{Projected 1-sigma error on $N_{\rm eff}$ as a function of the strength of the prior on the 6 cosmological parameters in the standard model. Each prior is expressed in units of the error that would be obtained internally from our suite of CMB experiments; e.g., $\sigma(H_0)_{old}=4.1\%$. The parameters that would be most useful to constrain externally for the purposes of determining $N_{\rm eff}$ are $n_s$ and $\Omega_bh^2$. The bottom-most curve shows the impact of imposing the given prior on {\it all} the other parameters simultaneously.}
\label{fig:prior_massless_neutrinos}
\end{center}
\end{figure}

\reffig{prior_massless_neutrinos} shows projections for how well our suite of CMB experiments will do at measuring $N_{\rm eff}$ as a function of priors on the other 6 parameters in the standard cosmological model ($H_0, \Omega_mh^2,\Omega_bh^2, n_s, A_s,$ and $\tau$). With no priors on the other parameters, the projected 1-sigma error is  about 0.02, suggesting a 2-sigma detection of the deviation from $N_{\rm eff}=3$. The sensitivity comes from the effect of extra species on the damping tail of the CMB anisotropies, both in temperature and polarization, so parameters that also affect the shape of the damping tail, like the slope $n_s$ and the baryon density $\Omega_bh^2$, are most degenerate with $N_{\rm eff}$. As such, \reffig{prior_massless_neutrinos} shows that obtaining external priors on either of these would reduce the errors on $N_{\rm eff}$ to ensure a 3-sigma detection of the partial decoupling prediction. The bottom-most curve shows that if all parameters were constrained externally, then the projected error on $N_{\rm eff}$ would improve significantly.

\subsection{Neutrino Masses}

Small scale structure formation is suppressed if fast-moving neutrinos comprise a significant part of the matter budget. This effect, probed by CMB lensing, will allow us to determine how much of the matter density is made up of neutrinos. Since the active neutrino number densities are known in the standard model, this can be directly transformed into a constraint on the the sum of the neutrino masses, $\sum m_\nu$. This is particularly exciting because there is a lower limit on $\sum m_\nu>0.05$ eV that emerges from oscillation experiments.

\begin{figure}[htbp]
\begin{center}
\includegraphics{prior_omnuh2_snow_mass.pdf}
\caption{Constraints on the sum of the neutrino masses as a function of priors on cosmological parameters, as in \reffig{prior_massless_neutrinos}.  Note that constraints would be improved significantly  by distance measurements that constrain $\Omega_ch^2$. The lower limit on the sum of the neutrino masses is 50 meV.}
\label{fig:prior_omeganuh2}
\end{center}
\end{figure}

\reffig{prior_omeganuh2} shows that, without any external priors, upcoming CMB experiments are poised to obtain 1-sigma limits on $\sum m_\nu$ of order 30 meV, corresponding to close to a 2-sigma detection even in the worst case scenario. Adding in external priors helps this significantly though, as discussed in \citet{2013arXiv1309.5383A}. There, the prior was deescribed as ``DESI BAO'', that is a measurement of distances as a function of redshift from the baryon acoustic oscillation feature probed by DESI. Here, the same improvement is parametrized by a strong prior on $\Omega_ch^2$, which indeed determines these low-redshift distances. As found in \cite{2013arXiv1309.5383A}, the projected error falls below 20 meV with this prior. Adding in other priors would help further.

@@

\subsection{Dark Energy Equation of State}

While the primordial CMB has limited information about the late-time dark energy equation of state, CMB lensing probes the growth of structure at late times and therefore is sensitive to $w$. Here we are using only the information from the lensing power spectrum as inferred from the CMB, but there is even more information not considered here contained in the many galaxy clusters that upcoming missions will detect via the Sunyaev-Zel'dovich effect.

\reffig{prior_w} shows that without any priors, CMB experiments will not do much better than current constraints, which hover around 10\%. However, an external constraint on the Hubble constant would improve the CMB constraints on $w$ considerably. {\bf We should try to understand why.}

{\bf I don't really understand this table.}
\begin{table}[htdp]

@@
